Appendix 1: Code Samples

### 2013-09-18 supp/code_samples.tex
Edits from Scott plus appendices updated

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[12pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

@@

\title{Code Samples}
%\author{The Author}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\section{\texttt{unmarked} code}

\subsection{Exponential model}
\subsection{Ricker-logistic model}
\subsection{Gompertz-logistic model}
\subsection{Exponential plus immigration model}
\subsection{Ricker-logistic plus immigration model}
\subsection{Gompertz-logistic plus immigration model}
\subsection{Autoregressive plus immigration model}
\subsection{Zero-inflated Poisson initial abundance model}
\subsection{Simulation study code}
\subsection{BBS analysis code}

@@

\subsection{Constant model}
\subsection{Exponential model}
\subsection{Ricker-logistic model}
\subsection{Gompertz-logistic model}
\subsection{Exponential plus immigration model}
\subsection{Ricker-logistic plus immigration model}
\subsection{Gompertz-logistic plus immigration model}
\subsection{Autoregressive plus immigration model}
\subsection{Zero-inflated Poisson dynamics model}
\subsection{Environmental stochasticity models}
\subsection{Forecast model}
\subsection{Simulation study code}
\subsection{BBS analysis code}

>>>>

\documentclass[12pt]{article}
\usepackage[a4paper,hmargin={2.4cm,2.4cm},vmargin={2.4cm,2.4cm}]{geometry}
\usepackage[mathlines,displaymath]{lineno}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[round,colon,authoryear]{natbib}
\usepackage{color, soul}
\usepackage{fancyvrb}
\usepackage{verbatim}
\usepackage[pdftex,colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,pdftex,pdfstartview=FitH,bookmarks=true]{hyperref}
\usepackage{rotating}

@@

\begin{document}
\title{Appendix 1: Code Samples}
\maketitle
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 
\section{\texttt{unmarked} code}

%\subsection{Exponential model}
%\subsection{Ricker-logistic model}
%\subsection{Gompertz-logistic model}
%\subsection{Exponential plus immigration model}
%\subsection{Ricker-logistic plus immigration model}
%\subsection{Gompertz-logistic plus immigration model}
%\subsection{Autoregressive plus immigration model}
%\subsection{Zero-inflated Poisson initial abundance model}
\subsection{Simulation study code}

File sim\_exp.R:
\begin{verbatim}
# Simulate data under exponential growth model
sim.exp <- function(lambda, r, p, nSites=100, nYears=45) {
    y <- N <- matrix(NA, nSites, nYears)
    N[,1] <- rpois(nSites, lambda)
    for(t in 2:nYears) {
        mu <- N[,t-1]*exp(r)
        N[,t] <- rpois(nSites, mu)
    }
    y[] <- rbinom(nSites*nYears, N, p)
    return(list(N=N, y=y, lambda=lambda, r=r, p=p,
                seed=.Random.seed, seed.kind=RNGkind()))
}
\end{verbatim}

File sim\_ricker.r:
\begin{verbatim}
# Simulate data under Ricker model
sim.ricker <- function(lambda, r, K, p, nSites=100, nYears=45) {
    y <- N <- matrix(NA, nSites, nYears)
    N[,1] <- rpois(nSites, lambda)
    for(t in 2:nYears) {
        mu <- N[,t-1]*exp(r*(1-N[,t-1]/K))
        N[,t] <- rpois(nSites, mu)
    }
    y[] <- rbinom(nSites*nYears, N, p)
    return(list(N=N, y=y, lambda=lambda, r=r, p=p,
                seed=.Random.seed, seed.kind=RNGkind()))
}
\end{verbatim}

File sim\_ricki.r:
\begin{verbatim}
# Simulate data under Ricker + immigration model
sim.ricki <- function(lambda, r, K, iota, p, nSites=100, nYears=45) {
    y <- N <- matrix(NA, nSites, nYears)
    N[,1] <- rpois(nSites, lambda)
    for(t in 2:nYears) {
        mu <- N[,t-1]*exp(r*(1-N[,t-1]/K)) + iota
        N[,t] <- rpois(nSites, mu)
    }
    y[] <- rbinom(nSites*nYears, N, p)
    return(list(N=N, y=y, lambda=lambda, r=r, K=K, iota=iota, p=p,
                seed=.Random.seed, seed.kind=RNGkind()))
}
\end{verbatim}

File ricker\_r0.1K5.r:
\begin{verbatim}
# Simulate data under Ricker model

# TO MODIFY
# (1) Change simulator... source(SOMETHING ELSE)
# (2) Change stub
# (3) Change simout
# (4) Change simulation parameters
# (5) Change pcountOpen arguments: "dynamics" and "K"
# (6) Results

ls()
stub <- "ricker_r0.1K5"  # (2)
setwd(paste('simulated', stub, sep='/'))
source("../sim_ricker.r")  # (1)
ls()

library(unmarked)

nsim <- 200
simout.mle <- simout.cover <- matrix(NA, nsim, 4)  # (3)
colnames(simout.mle) <- colnames(simout.cover) <- c("Lambda", "r", "K", "p") # (3)

# Simulate
set.seed(345489)
lambda <- 10 # (4)
r <- 0.1     # (4)
K <- 5       # (4)
p <- 0.25    # (4)
for(i in 1:nsim) {
    cat("\ndoing", i, "at", format(Sys.time()), "\n")
    data.name <- paste("data_", stub, "_", i, sep="")
    data.file <- paste("data/", data.name, ".gzip", sep="")
    sim.i <- sim.ricker(lambda=lambda, r=r, K=K, p=p, nSites=100, nYears=40) # (1)
    cat("   max(N)=", max(sim.i$N), "\n", sep="")
    assign(data.name, sim.i)
    save(list=data.name, file=data.file) # save simulated data
    umf.i <- unmarkedFramePCO(y=sim.i$y, numPrimary=ncol(sim.i$y))
    fm.i <- pcountOpen(~1, ~1, ~1, ~1, umf.i, K=200, dynamics="ricker",  # (5)
                       starts=c(log(lambda), log(r), log(K), qlogis(p))) # (5)
    model.name <- paste("fm_", stub, "_", i, sep="")
    model.file <- paste("fm/", model.name, ".gzip", sep="")
    assign(model.name, fm.i)
    save(list=model.name, file=model.file) # save fitted models
    mle.i <- coef(fm.i)
    simout.mle[i,] <- c(exp(mle.i[1]), exp(mle.i[2]), exp(mle.i[3]), plogis(mle.i[4])) # (6)
    cat("   mle=", round(simout.mle[i,], 4), "\n")
    write.csv(simout.mle, paste(stub, "_mle.csv", sep=""),
              row.names=FALSE)
    CI.lam <- confint(fm.i, type="lambda") # (6)
    CI.r <- confint(fm.i, type="gamma")    # (6)
    CI.K <- confint(fm.i, type="omega")    # (6)
    CI.p <- confint(fm.i, type="det")      # (6)
    simout.cover[i,1] <- log(lambda) >= CI.lam[1] & log(lambda) <= CI.lam[2] # (6)
    simout.cover[i,2] <- log(r) >= CI.r[1] & log(r) <= CI.r[2]               # (6)
    simout.cover[i,3] <- log(K) >= CI.K[1] & log(K) <= CI.K[2]               # (6)
    simout.cover[i,4] <- qlogis(p) >= CI.p[1] & qlogis(p) <= CI.p[2]         # (6)
    write.csv(simout.cover, paste(stub, "_cover.csv", sep=""),
              row.names=FALSE)
    rm(list=c(data.name, model.name))
}
\end{verbatim}

File ricki\_r0.05K10iota0.5.r:
\begin{verbatim}
# Simulate data under Ricker + immigration model

# TO MODIFY
# (1) Change simulator... source(SOMETHING ELSE)
# (2) Change stub
# (3) Change simout
# (4) Change simulation parameters
# (5) Change pcountOpen arguments: "dynamics" and "K"
# (6) Results

stub <- "ricki_r0.05K10iota0.5"  # (2)
setwd(paste('simulated', stub, sep='/'))
ls()
source("../sim_ricki.r")  # (1)
ls()

library(unmarked)

nsim <- 1000
simout.mle <- simout.cover <- matrix(NA, nsim, 5)  # (3)
colnames(simout.mle) <- colnames(simout.cover) <-
    c("Lambda", "r", "K", "iota", "p") # (3)

# Simulate
set.seed(345489)
lambda <- 10 # (4)
r <- 0.05    # (4)
K <- 10      # (4)
iota <- 0.5  # (4)
p <- 0.25    # (4)
for(i in 1:nsim) {
    cat("\ndoing", i, "at", format(Sys.time()), "\n")
    data.name <- paste("data_", stub, "_", i, sep="")
    data.file <- paste("data/", data.name, ".gzip", sep="")
    # Create simulated data
    sim.i <- sim.ricki(lambda=lambda, r=r, K=K, iota=iota, p=p, nSites=100, nYears=40) # (1)
    cat("   max(N)=", max(sim.i$N), "\n", sep="")
    assign(data.name, sim.i)
    # Save simulated data
    save(list=data.name, file=data.file)
    # Run pcountOpen model 
    umf.i <- unmarkedFramePCO(y=sim.i$y, numPrimary=ncol(sim.i$y))
    fm.i <- pcountOpen(~1, ~1, ~1, ~1, umf.i, K=200, dynamics="ricker", # (5)
      immigration=TRUE, starts=c(log(lambda), log(r), log(K), qlogis(p), log(iota))) # (5)
    # Save fitted models
    model.name <- paste("fm_", stub, "_", i, sep="")
    model.file <- paste("fm/", model.name, ".gzip", sep="")
    assign(model.name, fm.i)
    save(list=model.name, file=model.file) 
    # Save maximum likelihood estimates
    mle.i <- coef(fm.i)
    simout.mle[i,] <- c(exp(mle.i[1]), exp(mle.i[2]), exp(mle.i[3]), exp(mle.i[5]), plogis(mle.i[4])) # (6)
    cat("   mle=", round(simout.mle[i,], 4), "\n")
    write.csv(simout.mle, paste(stub, "_mle.csv", sep=""),
              row.names=FALSE)
    # Save coverage
    CI.lam <- confint(fm.i, type="lambda") # (6)
    CI.r <- confint(fm.i, type="gamma")    # (6)
    CI.K <- confint(fm.i, type="omega")    # (6)
    CI.p <- confint(fm.i, type="det")      # (6)
    CI.iota <- confint(fm.i, type="iota")  # (6)
    simout.cover[i,1] <- log(lambda) >= CI.lam[1] & log(lambda) <= CI.lam[2] # (6)
    simout.cover[i,2] <- log(r) >= CI.r[1] & log(r) <= CI.r[2]               # (6)
    simout.cover[i,3] <- log(K) >= CI.K[1] & log(K) <= CI.K[2]               # (6)
    simout.cover[i,4] <- log(iota) >= CI.iota[1] & log(iota) <= CI.iota[2]   # (6)
    simout.cover[i,5] <- qlogis(p) >= CI.p[1] & qlogis(p) <= CI.p[2]         # (6)
    write.csv(simout.cover, paste(stub, "_cover.csv", sep=""),
              row.names=FALSE)
    # Clean up
    rm(list=c(data.name, model.name))
}
\end{verbatim}

\subsection{BBS analysis code}
\begin{verbatim}
library(unmarked)

# Read in input files
ydata <- as.matrix(read.csv('oven3a.csv', row.names=1))
observers <- as.matrix(read.csv('observers.csv', row.names=1))
first.run <- as.matrix(read.csv('first.run.csv', row.names=1))
wind.factor <- as.matrix(read.csv('wind.factor.csv', row.names=1))
# Fix formatting
wind.factor[wind.factor==" 0"] <- "0"
wind.factor[wind.factor==" 1"] <- "1"
wind.factor[wind.factor==" 2"] <- "2"
# Combine obsCovs together into list
obs4 <- list(observers = observers, first.run = first.run, wind.factor = 
  wind.factor)

# Basic settings and labels
year1 <- 66
aou <- 'oven'
gam.model <- list(~1)[[1]] 
om.model <- list(~1)[[1]] 
lam.model <- list(~1)[[1]] 
K <- 600
iota.model <- list(~1) [[1]]

# Input data for pcountOpen
bird.frame <- unmarkedFramePCO(ydata, obsCovs=obs4, numPrimary=45)

# Run mixture models for initial abundance  
mods1 <- rep('', 3)
outputs1 = list()
for (i in 1:3) {
  mixture <- c('P','NB','ZIP')[i]
  dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[3]
  p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[1]]
  immigration <- FALSE
  mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
    paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
    substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
    'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
    ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
    'K', sep='.')))), K, sep="")
  mods1[i] <- mod.name
  file.name <- paste(mod.name, 'gzip', sep='.')
  print(file.name)
  fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
    lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
    dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
  outputs1[[i]] <- fm
  print(summary(fm))
  save(fm, file=file.name)
}
names(outputs1) <- mods1

AIC.table <- function(mods, sort=T) {
  tab <- data.frame(model = names(mods), npar = NA, AIC = NA, delta = NA, 
    weight = NA)
  for (i in 1:length(mods)) {
    tab$npar[i] <- length(coef(mods[[i]]))
    tab$AIC[i] <- mods[[i]]@AIC
  }
  tab$delta <- tab$AIC - min(tab$AIC)
  tab$weight <- exp(-tab$delta/2)/sum(exp(-tab$delta/2))
  if (sort)
    tab <- tab[order(tab$AIC), ]
  return(tab)
}

# Compare these models
mix.aic <- AIC.table(outputs1)
mix.aic
mix.num <- as.integer(rownames(mix.aic)[1])
mix.num

# Run models for detection probability  
mods2 <- c(mods1[mix.num], rep('', 3))
outputs2 <- list(outputs1[[mix.num]])
for (i in 2:4) {
  mixture <- c('P','NB','ZIP')[mix.num]
  dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[3]
  p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[i]]
  immigration <- FALSE
  mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
    paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
    substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
    'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
    ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
    'K', sep='.')))), K, sep="")
  mods2[i] <- mod.name
  file.name <- paste(mod.name, 'gzip', sep='.')
  print(file.name)
  fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
    lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
    dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
  outputs2[[i]] <- fm
  print(summary(fm))
  save(fm, file=file.name)
}
names(outputs2) <- mods2

# Compare these models
p.aic <- AIC.table(outputs2)
p.aic
p.num <- as.integer(rownames(p.aic)[1])
p.num

# Run models for population dynamics  
mods3 <- c(mods2[p.num], rep('', 8))
outputs3 <- list(outputs2[[p.num]])
# Constant dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[1]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[2] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[2]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Autoregressive dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[2]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[3] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[3]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Ricker dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[4]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[4] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, 
# so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:length(starts)])
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model, 
  starts=starts)
outputs3[[4]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Gompertz dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[5]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[5] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:length(starts)])
fm <- pcountOpen(data=bird.frame, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  mixture=mixture, dynamics=dynamics, immigration=immigration, 
  iotaformula=iota.model, starts=starts)
outputs3[[5]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Trend + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[3]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[6] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[6]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Autoregressive + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[2]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[7] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[7]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Ricker + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[4]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[8] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, 
# so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:(length(starts)-1)], 
  -1.5, starts[length(starts)])
fm <- pcountOpen(data=bird.frame, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  mixture=mixture, dynamics=dynamics, immigration=immigration, 
  iotaformula=iota.model, starts=starts)
outputs3[[8]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Gompertz + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[5]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[9] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, 
# so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:(length(starts)-1)], 
  -1.5, starts[length(starts)])
fm <- pcountOpen(data=bird.frame, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  mixture=mixture, dynamics=dynamics, immigration=immigration, 
  iotaformula=iota.model, starts=starts)
outputs3[[9]] <- fm
print(summary(fm))
save(fm, file=file.name)

names(outputs3) <- mods3

# Compare these models
dyn.aic <- AIC.table(outputs3)
dyn.aic
dyn.num <- as.integer(rownames(dyn.aic)[1])
dyn.num
\end{verbatim}

@@

\subsection{Geometric-recruitment (autoregressive) model}
\begin{verbatim}
model {
lambda ~ dunif(0, 5)
omega ~ dunif(0,1)
gamma ~ dunif(0, 10)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    S[i,t-1] ~ dbin(omega, N[i,t-1])
    G[i,t-1] ~ dpois(gamma * N[i,t-1])
    N[i,t] <- S[i,t-1] + G[i,t-1]
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Exponential model}
\begin{verbatim}
model {
lambda ~ dunif(0, 5)
r ~ dnorm(0,0.01)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    N[i,t] ~ dpois(exp(r) * N[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Ricker-logistic model}
\begin{verbatim}
model {
lambda ~ dunif(0, 50)
r ~ dunif(0,5)
K ~ dunif(0, 50)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    muN[i,t-1] <- N[i,t-1] * exp(r * (1 - N[i,t-1] / K))
    N[i,t] ~ dpois(muN[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Gompertz-logistic model}
\begin{verbatim}
model {
lambda ~ dunif(0, 50)
r ~ dunif(0,5)
K ~ dunif(0, 50)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    muN[i,t-1] <- N[i,t-1] * exp(r * (1 - log(N[i,t-1] + 1) / log(K + 1)))
    N[i,t] ~ dpois(muN[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Exponential plus immigration model}
\begin{verbatim}
model {
lambda ~ dunif(0, 5)
r ~ dnorm(0,0.01)
p ~ dunif(0,1)
iota ~ dunif(0, 5)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    muN[i,t-1] <- N[i,t-1] * exp(r) + iota
    N[i,t] ~ dpois(muN[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Zero-inflated Poisson dynamics model}
\begin{verbatim}
model {
lambda ~ dunif(0, 200)
psi ~ dunif(0, 1)
# Trend + immigration
r ~ dnorm(0, 0.01)
iota ~ dunif(0, 20)
p ~ dunif(0,1)

for(i in 1:nSites) {
  range[i] ~ dbern(1-psi)
  N[i,1] ~ dpois(lambda*range[i])
  for(t in 2:nYears) {
      muN[i,t-1] <- (N[i,t-1]*exp(r)+iota)*range[i]
      N[i,t] ~ dpois(muN[i,t-1])
    }
  for(t in 1:nYears) {
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Environmental stochasticity models}
\begin{verbatim}
model {
# Explicitly modeling dynamic parameters as functions of environmental covariates
# Negative binomial initial abundance, Gompertz-logistic dynamics
lambda ~ dunif(0, 400)
alpha ~ dunif(0, 400)
P <- alpha/(alpha + lambda)
K0 ~ dnorm(0, 0.01)
Krain ~ dnorm(0, 0.01)
r0 ~ dnorm(0, 0.01)
rTemp ~ dnorm(0, 0.01)
p ~ dunif(0,1)

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(P, alpha)
  for(t in 2:nYears) {
      K[i,t-1] <- exp(K0 + Krain*rain[i,t-1]) 
      r[i,t-1] <- exp(r0 + rTemp*temp[i,t-1]) 
      muN[i,t-1] <- N[i,t-1] * exp(r[i,t-1] * (1 - log(N[i,t-1] + 1) / log(K[i,t-1] + 1)))
      N[i,t] ~ dpois(muN[i,t-1])
  }
  for(t in 1:nYears) {
    y[i,t] ~ dbin(p, N[i,t])
  }
}
}
\end{verbatim}

\begin{verbatim}
model {
# Independent environmental stochasticity between sites
# Negative binomial initial abundance, Ricker-logistic dynamics + immigration
lambda ~ dunif(0, 400)
alpha ~ dunif(0, 400)
P <- alpha/(alpha + lambda)
K ~ dunif(0, 400)
r ~ dunif(0, 5)
iota ~ dunif(0, 10)
sigma.nu ~ dunif(0, 2)
tau.nu <- 1 / (sigma.nu * sigma.nu)
p ~ dunif(0, 1)

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(P, alpha)
  for(t in 2:nYears) {
      nu[i,t-1] ~ dnorm(0, tau.nu) 
      muN[i,t-1] <- N[i, t-1] * exp(nu[i, t-1] + r * (1 - log(N[i, t-1] + 1) / log(K + 1))) + iota
      N[i,t] ~ dpois(muN[i, t-1])
  }
  for(t in 1:nYears) {
    y[i,t] ~ dbin(p, N[i,t])
  }
}
}
\end{verbatim}
